Mathematical Aspects of the Capital Market

### 2017-08-11 phd/Chapterintroduction.tex
Thomas is not happy about the contribution part.

Now consider $p$ return series
$X_{i,t}$, $i=1,2,\dots, p; t=1,\dots,n$.
Suppose each of these series is a linear combination of $K$ factors
$Y_{j,t}$, $j=1,2,\dots,K$, the $j$-th of which is regularly varying
with index $\alpha_j$. Then by the summation property, each and every
$\{X_i\}$ is regularly varying with index $\min_{1 \leq j \leq K} \alpha_j$.
In practice a factor $Y_{i,t}$ is estimated as
$\hat Y_{j,t} = \sum_{i=1}^p a_{j, i} X_{i,t}$, where
$(a_{j, 1}, a_{j, 2}, \dots, a_{j, p})^\top$ is the $j$-th eigenvector
of the sample covariance matrix of
$\{X_{i,t}\}, i=1,\dots,p; t=1,\dots,n$, i.e. the eigenvector
associated with the $j$-th largest eigenvalue. 
For this reason, it is important to understand the eigensystem of the
sample covariance matrix of $\{X_{i,t}\}$. This topic is addressed in
chapters \ref{ch:bernoulli} and \ref{ch:extremes} of this thesis.

@@

\section{Stochastic Recurrence Equation}
One of the most important dynamical mechanisms that lead to regularly
varying r.v. is stochastic recursion of the following form:
\begin{equation}
  \label{eq:rhjyu}
  X_t = A_t X_{t-1} + B_t, \quad t \in \mathbb Z
\end{equation}
where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
random matrix and $B_t$ is a $d$-dimensional vector, random or
deterministic. The sequence $\{(A_t, B_t)\}_{t \in \mathbb Z}$ are
iid. The stationary solution to \eqref{eq:rhjyu} satisfies the fixed-point
equation $X \overset{d}{=} A X + B$, where $X$ and $(A, B)$ are
generic elements of the $\{X_t\}_{t \in \mathbb Z}$ and
$(A_t, B_t)_{t \in \mathbb Z}$ sequences.

Kesten \cite{kesten:1973} showed that, when $A_t$ are almost
surely non-negative, has no row or column of only zeros, and
$B_t$ is almost surely non-negative and is not equal to the null
vector with probability 1, then the solution $X$ to the equation
$X \overset{d}{=} A X + B$ is regularly varying with a positive index
$\alpha$, assuming the following conditions (M) and (A):
\begin{itemize}
\item Condition (M)
  \begin{enumerate}
  \item The top Lyapunov exponent
    \[
    \gamma = \inf_{n \geq 1} {1 \over n}\E \log \|A_n \cdots A_1\|
    \]
    is negative.
  \item There exists $\alpha > 0$ such that
    \[
    1 = \lambda(\alpha) = \lim_{n \to \infty} {1 \over n} \log \E \|A_n \cdots A_1\|^\alpha
    \]
  \item $\E (\|A_1\|^\alpha \log^+\|A_1\|) < \infty$
  \item $\E |B_1|^\alpha < \infty$
  \end{enumerate}
\item Condition (A) : The group generated by
  \[
  \{\log\rho(s): s = A_n \cdots A_1 \text{ for some } n \geq 1\}
  \]
  is dense in $\reals$, where $\rho(s)$ denotes the spectral
  radius of matrix $s$.
\end{itemize}
Upon these conditions, Kesten's theorem gives
\begin{equation}
  \label{eq:jorg}
  u^\alpha \P(u^{-1} V \in \cdot) \overset{v}{\to} \nu_\alpha(\cdot)
\end{equation}
where $\nu_\alpha$ is a non-null Radon measure on
$\reals^d_+ \setminus \{0\}$ with the property
$\nu_\alpha(a A) = a^{-\alpha} \nu_\alpha(A)$.

In addition to non-negative matrices, two other classes of random
matrices have been shown to lead to power-law tails via the recurrence
relation \eqref{eq:garchpq_sre}. Alsmeyer and Mentemeier
\cite{alsmeyer:mentemeier:2012} considered invertible matrices whose
distribution has a density. Let $M(d, \reals)$ denote the space of
$d \times d$ matrices with real entries that are invertible with
probability 1. They replaced Kesten's condition
$\E (\|A\|^\alpha \log^+\|A\|) < \infty$ with a stronger
counterpart
$\E [\|A\|^\alpha (\log^+\|A\| + \log \|A^{-1} \|)] < \infty$,
and lifted the condition (A). In addition, they assumed
\begin{enumerate}
  \item The Markov chain $X_n$ on $\sphere^{d-1}$, namely
    $X_n = A_n X_{n-1} / |A_n X_{n-1}|$, is irreducible, i.e. for any open
    set $U \subset \sphere^{d-1}$ and any $u \in \sphere^{d-1}$, 
    $\exists n \geq 1$ such that $\P(X_n u \in U) > 0$.
  \item There exist $N \geq 1$, $c, \epsilon > 0$ and an invertible
    matrix $\bar A \in M(d, \reals)$ such that for any set
    $C \subset M(d, \reals)$, it holds true
    $\P(A_N \cdots A_1 \in C) \geq c |B_\epsilon(\bar A) \cap C|$,
    where $|\cdot|$ denotes the Lebesgue measure.
\end{enumerate}
These assumptions are termed conditions (id). Furthermore, they assumed
that there was no point in $\reals^d$ such that the recurrence
equation \eqref{eq:garchpq_sre} was stuck at this point with probability 1: 
$\P(A X + B = X) < 1$ for all $X \in \reals^d$. With these
assumptions, they showed
\[
\lim_{u \to \infty} u^\alpha \P(\inn{x, V} > u) = e_\alpha(x)
\]
where $x \in \sphere^{d - 1}$ and $e_\alpha(\cdot)$ is a continuous
function on $\sphere^{d-1}$.

@@

Bollerslev \cite{bollerslev:1986} showed that the fixed-point equation
$Y \overset{d}{=} A Y + B$ has a unique, strictly stationary solution
$Y$ with finite variance if and only if
\begin{equation}
  \sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1
  \label{eq:bollerslev:intro}
\end{equation}

\section{GARCH models}
Introduced by Bollerslev \cite{bollerslev:1986} in 1986,
{\em Generalized Autoregressive Conditional Heteroscedasticity}
(GARCH) models have been hugely popular for modelling volatility of
financial time series and have inspired numerous variants.
A GARCH($p, q$) model is a stationary process
$\{X_t\}_{t \in \mathbb Z}$satisfying
\begin{eqnarray*}
  X_t &=& \sigma_t Z_t \\
  \sigma_t^2 &=& \omega + \sum_{i=1}^p \alpha_i X_{t-i}^2 +
  \sum_{j=1}^q \beta_j \sigma_{t-j}^2
\end{eqnarray*}
where $\{X_t\}$ is a return series, e.g. stock returns, foreign exchange
rates, interest rates, etc; $\{Z_t\}$ is an iid, mean 0, unit-variance
sequence, $\sigma_t^2$ is the variance of the distribution of $X_t$
conditional on
$\{(X_i, \sigma_i^2)\}_{i=0}^{t-1}$; $\omega, \{\alpha_i\}_{i=1}^p,
\{\beta_i\}_{i=1}^q$
are non-negative parameters of the model. The GARCH($p,q$) recurrence
equation is of the form of \eqref{eq:rhjyu}. With appropriate
conditions, $\sigma_t^2$ is shown to be a positive Harris recurrent
Markov chain (cf. Bollerslev \cite{bollerslev:1986} and Buraczewski et al.
\cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
has regularly varying tails. The tail index $\alpha$ is given by
\begin{equation}
  \label{eq:grgbg}
  \lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\alpha = 0,
\end{equation}
where $\{A_i\}_{i \in \mathbb Z}$ are iid matrices whose entries are
functions of $\{\alpha_i\}_{i=1}^p$, $\{\beta_i\}_{i=1}^q$ and
$\{Z_t^2\}$:
\[
A_i =
\begin{pmatrix}
  \alpha_1 Z_{t-1}^2 + \beta_1 & \beta_2 & \cdots &
  \beta_{q-1} & \beta_q & \alpha_2 & \alpha_3 &
  \cdots & \alpha_{p-1} & \alpha_p\\
  1 & 0 & \cdots & 
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots & 
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  1 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  Z_{t-1}^2 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  0 & 0 & 1 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots &
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\    
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 1 & 0 \\    
\end{pmatrix}
\]
While GARCH models have been very successful for modelling financial
time series, it does have its drawbacks. For example, the tail
index is very sensitive to the model parameters $\{\alpha_i\}_{i=1}^p$ and
$\{\beta_i\}_{i=1}^q$. In applications, these parameters need to be
estimated from a sample and are always uncertain to some extent. For
this reason, there can be a significant discrepancy between the tail
index obtained via \eqref{eq:grgbg} with the model parameters
substituted for their sample estimates and the Hill estimate
\eqref{eq:fhhty}.

@@

In the pioneering work of Clark \cite{clark:1973}, the author modelled
the logarithmic price $Y_t$ as a subordinated stochastic process:
$Y_t = V_{\tau_t}$, where $V_i$ is a normally distributed random
variable with mean 0, variance $\sigma^2 i$ and independent
increments. $\tau_t$ is a real-valued, non-negative, non-decreasing
sequence with $\tau_0 = 0$. It models a time change. As pointed out by
Shepard and Andersen \cite{Shephard:Andersen:2009}, the log-price
process $Y_t$ is serially uncorrelated although potentially dependent,
provided that $V_t$ and $\tau_t$ are independent.

Later authors, e.g. Back \cite{back:1991} chose to model the log-price
process as a semimartingale, with increments of the martingale
component modelled as a product process:
\begin{eqnarray}
  Y_t &=& Y_0 + A_t + M_t \nonumber \\
  M_t - M_{t-1} &=& X_t = \sigma_t Z_t
  \label{eq:rtght}
\end{eqnarray}
where $A_t$ is a finite variation process and $M_t$ is a martingale
and hence $Y_t$ is a semimartingale. $\sigma_t$ is non-negative and
$Z_t$ is an iid process with 0 mean and unit variance. A convenient
choice of $\sigma_t$ is
\begin{equation}
  \label{eq:rfht}
  \log\sigma_t = \sum_{l \in \mathbb Z} \psi_l \eta_{t-l}  
\end{equation}
where $-1 \leq \{\psi_l\}_{l \in \mathbb Z} \leq 1$
is a sequence of real numbers with at least one non-zero element,
$\{\eta_t\}_{t \in \mathbb Z}$ is a iid sequence of random
variables. By Kolmogorov's 3-series theorem, the infinite series above
converges if and only if $\sum_{l \in \mathbb Z} \psi_l^2 < \infty$.
In particular, if $\{\eta_i\}_{i \in \mathbb Z}$ is normally
distributed with 0 mean, $\{\sigma_t\}_{t \in \mathbb Z}$ is
stationary and its marginal distributions are log-normal.

If, however, the sequence $\{Z_t\}_{t \in \mathbb Z}$ or
$\{\sigma_t\}_{t \in \mathbb Z}$ is regularly varying with index
$\alpha$ and some additional conditions are satisfied, $X_t$ is also
regularly varying with the same index. Specifically, if $Z_t$ is
regularly varying and $\sigma_t$ has a lighter tail, the conclusion
follows from Breiman's lemma. See \S\ref{sec:case1} of Jan\ss en et
al. i.e. chapter \ref{ch:bernoulli} of this thesis for more details.

If, instead, $\sigma_t$ has a heavier tail than $Z_t$, then $X_{i,t}$
is regularly varying assuming the following conditions:
\begin{itemize}
\item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
  is a generic element of the iid sequence
  $\{\eta_t\}_{t \in \mathbb Z}$, and
  \begin{equation}
    \label{eq:dfhty}
    \E e^{\alpha \eta} = \infty, \quad\text{ or }
    {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
  \end{equation}
    where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
    iid sequence $\{\eta_t\}_{t \in \mathbb Z}$.
\end{itemize}
Note the only finite limit for
${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ is
$\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
these two conditions lead to rather different behavior for
eigenvalues of the sample covariance matrix of stochastic volatility
processes of the form \eqref{eq:rfht}.

There are a few advantages in using stochastic volatility models. For
one thing, they are among the simplest models allowing for conditional
heteroscedasticity (cf. Andersen et
al. \cite{andersen:davis:kreiss:mikosch:2009}); for another thing,
they greatly improve the accuracy of predicting future volatilities
by taking advantage of high frequency data. Specifically, for the
semimartingale $Y_t$ we have
\begin{equation}
  \label{eq:jieg}
  [Y_{\delta}]_t
  =
  \sum_{j=1}^{\floor{t/\delta}}
  (Y_{j \delta} - Y_{(j-1)\delta})^2
  \overset{P}{\to} [Y]_t,
  \quad
  \delta \to 0.
\end{equation}
The object $[Y_\delta]_t$ is called the realized quadratic variation
in time series literature.
The convergence in probability follows directly from the definition of
quadratic variation.
%% When $\E Z_t = 0$, $[Y]_t = [M]_t$ 
Furthermore, Jacod \cite{jacod:1994} and
Barndorff-Nielsen and Shephard \cite{barndorff:shephard:2002} proved a
central limit theorem:
\[
  {[Y_\delta]_t - [Y]_t \over \sqrt{2 \delta \int_0^t \sigma_s^2 ds}}
  \overset{d}{\to} N(0, 1),
  \quad \delta \to 0
\]
Moreover, the structure $X_t = M_{t} - M_{t-1} = \sigma_t Z_t$ gives
$[Y_\delta]_t \overset{P}{\to} \int_0^t \sigma_s^2 ds$. Meanwhile,
It\^o's formula gives, for semimartingale $Y_t$,
\begin{eqnarray}
  Y_t^2 &=& [Y]_t + 2 \int_0^t Y_s d Y_s \nonumber \\
  &=& [Y]_t + 2 \int_0^t Y_s dA_s + 2 \int_0^t Y_s dM_s \nonumber \\
  \E Y_t^2 &=& \E [Y]_t + 2 \E \left( \int_0^t Y_s dA_s \right) \nonumber\\
  &\approx& \E [Y]_t, \quad \text{when $t$ is small} \label{eq:hytre}
\end{eqnarray}
\eqref{eq:hytre} and \eqref{eq:jieg} shows $\E Y_t^2$, or in other
words, the forecast of future squared return can be obtained as
$\E [Y_\delta]_t$, i.e. the forecast of future realized quadratic
variation. Since past realized quadratic variations are observable,
forcast of future $[Y_\delta]_t$ can be obtained using e.g. an ARFIMA
model.

@@

On the other hand, we have also seen that, when the left and the right
tails have the same indices, investor preference over the equity has
more sophisticated variations in the parameters' space including the
tail parameters of the equity, the interest rate, the investor's risk
apetite as captured by his utility function, and his threshold of
disappointment.

>>>>

Now consider $p$ return series
$X_{i,t}$, $i=1,2,\dots, p; t=1,\dots,n$.
Suppose each of these series is a linear combination of $K$ factors
$Y_{j,t}$, $j=1,2,\dots,K$, the $j$-th of which is regularly varying
with index $\alpha_j$. Then by the summation property, each and every
$\{X_{i,t}\}$ is regularly varying with index $\min_{1 \leq j \leq K} \alpha_j$.
In practice a factor $Y_{i,t}$ is estimated as
$\hat Y_{j,t} = \sum_{i=1}^p a_{j, i} X_{i,t}$, where
$(a_{j, 1}, a_{j, 2}, \dots, a_{j, p})^\top$ is the $j$-th eigenvector
of the sample covariance matrix of
$\{X_{i,t}\}, i=1,\dots,p; t=1,\dots,n$, i.e. the eigenvector
associated with the $j$-th largest eigenvalue. 
For this reason, it is important to understand the eigensystem of the
sample covariance matrix of $\{X_{i,t}\}$. This topic is addressed in
chapters \ref{ch:bernoulli} and \ref{ch:extremes} of this thesis.

@@

\section{Stochastic Recurrence Equation}
One of the most important dynamical mechanisms that lead to regularly
varying random variables is a stochastic recursion of the following form:
\begin{equation}
  \label{eq:rhjyu}
  X_t = A_t X_{t-1} + B_t, \quad t \in \mathbb Z
\end{equation}
where $X_t$ is a $d$-dimensional random vector, $A_t$ is a $d\times d$
random matrix and $B_t$ is a $d$-dimensional vector, random or
deterministic. The sequence $\{(A_t, B_t)\}_{t \in \mathbb Z}$ is
iid. The stationary solution to \eqref{eq:rhjyu} satisfies the fixed-point
equation $X \overset{d}{=} A X + B$, where $X$ and $(A, B)$ are
generic elements of the $\{X_t\}_{t \in \mathbb Z}$ and
$(A_t, B_t)_{t \in \mathbb Z}$ sequences.

Kesten \cite{kesten:1973} showed that, when $A_t$ are almost
surely non-negative, has no row or column of only zeros, and
$B_t$ is almost surely non-negative and is not equal to the null
vector with probability 1, then the solution $X$ to the equation
$X \overset{d}{=} A X + B$ is regularly varying with a positive index
$\alpha$, assuming the following conditions (M) and (A):
\begin{itemize}
\item Condition (M)
  \begin{enumerate}
  \item The top Lyapunov exponent
    \[
    \gamma = \inf_{n \geq 1} {1 \over n}\E \log \|A_n \cdots A_1\|
    \]
    is negative.
  \item There exists $\alpha > 0$ such that
    \[
    1 = \lambda(\alpha) = \lim_{n \to \infty} {1 \over n} \log \E \|A_n \cdots A_1\|^\alpha
    \]
  \item $\E (\|A_1\|^\alpha \log^+\|A_1\|) < \infty$
  \item $\E |B_1|^\alpha < \infty$
  \end{enumerate}
\item Condition (A) : The group generated by
  \[
  \{\log\rho(s): s = A_n \cdots A_1 \text{ for some } n \geq 1\}
  \]
  is dense in $\reals$, where $\rho(s)$ denotes the spectral
  radius of matrix $s$.
\end{itemize}
Upon these conditions, Kesten's theorem gives
\begin{equation}
  \label{eq:jorg}
  u^\alpha \P(u^{-1} X \in \cdot) \overset{v}{\to} \nu_\alpha(\cdot)
\end{equation}
where $\nu_\alpha$ is a non-null Radon measure on
$\reals^d_+ \setminus \{0\}$ with the property
$\forall a > 0, \nu_\alpha(a A) = a^{-\alpha} \nu_\alpha(A)$.

In addition to non-negative matrices, two other classes of random
matrices have been shown to lead to power-law tails via the recurrence
relation \eqref{eq:garchpq_sre}. Alsmeyer and Mentemeier
\cite{alsmeyer:mentemeier:2012} considered invertible matrices whose
distribution has a density. Let $M(d, \reals)$ denote the space of
$d \times d$ matrices with real entries that are invertible with
probability 1. They replaced Kesten's condition
$\E (\|A\|^\alpha \log^+\|A\|) < \infty$ with a stronger
counterpart
$\E [\|A\|^\alpha (\log^+\|A\| + \log \|A^{-1} \|)] < \infty$,
and lifted the condition (A). In addition, they assumed
\begin{enumerate}
  \item For any open set $U \subset \sphere^{d-1}$ and any
    $u \in \sphere^{d-1}$,  $\exists n \geq 1$ such that
    \[
    \P\left({\prod_{i=1} A_i u
        \over
        \left|\prod_{i=1} A_i u\right|}\in U
    \right) > 0.
    \]
  \item There exist $N \geq 1$, $c, \epsilon > 0$ and an invertible
    matrix $\bar A \in M(d, \reals)$ such that for any set
    $C \subset M(d, \reals)$, it holds true
    $\P(A_N \cdots A_1 \in C) \geq c |B_\epsilon(\bar A) \cap C|$,
    where $|\cdot|$ denotes the Lebesgue measure.
\end{enumerate}
These assumptions are termed conditions (id). Furthermore, they assumed
that there was no point in $\reals^d$ such that the recurrence
equation \eqref{eq:garchpq_sre} was stuck at this point with probability 1: 
$\P(A X + B = X) < 1$ for all $X \in \reals^d$. With these
assumptions, they showed
\[
\lim_{u \to \infty} u^\alpha \P(\inn{x, X} > u) = e_\alpha(x)
\]
where $x \in \sphere^{d - 1}$ and $e_\alpha(\cdot)$ is a continuous
function on $\sphere^{d-1}$.

@@

% Bollerslev \cite{bollerslev:1986} showed that the fixed-point equation
% $Y \overset{d}{=} A Y + B$ has a unique, strictly stationary solution
% $Y$ with finite variance if and only if
% \begin{equation}
%   \sum_{i=1}^p \alpha_i + \sum_{j=1}^q \beta_j < 1
%   \label{eq:bollerslev:intro}
% \end{equation}

\section{GARCH models}
Introduced by Bollerslev \cite{bollerslev:1986} in 1986,
{\em Generalized Autoregressive Conditional Heteroscedasticity}
(GARCH) models have been hugely popular for modelling volatility of
financial time series and have inspired numerous variants.
A GARCH($p, q$) model is a stationary process
$\{X_t\}_{t \in \mathbb Z}$satisfying
\begin{eqnarray*}
  X_t &=& \sigma_t Z_t \\
  \sigma_t^2 &=& \omega + \sum_{i=1}^p \alpha_i X_{t-i}^2 +
  \sum_{j=1}^q \beta_j \sigma_{t-j}^2
\end{eqnarray*}
where $\{X_t\}$ is a return series, e.g. stock returns, foreign exchange
rates, interest rates, etc; $\{Z_t\}$ is an iid, mean 0, unit-variance
sequence, $\sigma_t^2$ is the variance of the distribution of $X_t$
conditional on
$\{(X_i, \sigma_i^2)\}_{i=0}^{t-1}$; $\omega, \{\alpha_i\}_{i=1}^p,
\{\beta_i\}_{i=1}^q$ are non-negative parameters of the model.
Written in matrix form, as shown in \eqref{eq:jopger}, the
GARCH($p,q$) recurrence  equation is of the form of
\eqref{eq:rhjyu}. With appropriate conditions,
$(\sigma_t^2, \dots, \sigma_{t-q+1}^2, X_{t-1}^2, \dots, X_{t - p
  +1}^2)^\top$ is shown to be a positive Harris recurrent Markov chain
(cf. Bollerslev \cite{bollerslev:1986} and Buraczewski et al.
\cite{buraczewski:damek:mikosch:2016}), whose stationary distribution
has regularly varying tails. The tail index $\alpha$ is given by
\begin{equation}
  \label{eq:grgbg}
  \lim_{n \to \infty} {1 \over n}\log\E\|A_n \cdots A_1\|^\alpha = 0,
\end{equation}
where $\{A_i\}_{i \in \mathbb Z}$ are iid matrices whose entries are
functions of $\{\alpha_i\}_{i=1}^p$, $\{\beta_i\}_{i=1}^q$ and
$\{Z_t^2\}$:
\[
A_i =
\begin{pmatrix}
  \alpha_1 Z_{t-1}^2 + \beta_1 & \beta_2 & \cdots &
  \beta_{q-1} & \beta_q & \alpha_2 & \alpha_3 &
  \cdots & \alpha_{p-1} & \alpha_p\\
  1 & 0 & \cdots & 
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots & 
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  1 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  Z_{t-1}^2 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\
  0 & 0 & \cdots &
  0 & 0 & 1 & 0 & \cdots & 0 & 0 \\
  \vdots & \vdots & \ddots &
  \vdots & \vdots & \vdots & \vdots &
  \ddots & \vdots & \vdots \\
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 0 & 0 \\    
  0 & 0 & \cdots &
  0 & 0 & 0 & 0 & \cdots & 1 & 0 \\    
\end{pmatrix}
\]
While GARCH models have been very successful for modelling financial
time series, it does have its drawbacks. For example, the tail
index is very sensitive to the model parameters $\{\alpha_i\}_{i=1}^p$ and
$\{\beta_i\}_{i=1}^q$. In applications, these parameters need to be
estimated from a sample and are always uncertain to some extent. For
this reason, there can be a significant discrepancy between the tail
index obtained via \eqref{eq:grgbg} with the model parameters
substituted for their sample estimates and the Hill estimate
\eqref{eq:fhhty}.

@@

In the pioneering work of Clark \cite{clark:1973}, the author modelled
the logarithmic price $Y_t$ as a subordinated stochastic process:
$Y_t = V_{\tau_t}$, where $V_i/\sigma^2$ is a standard Brownian
motion. $\tau_t$ is a real-valued, non-negative, non-decreasing
sequence with $\tau_0 = 0$. It models a time change. As pointed out by
Shepard and Andersen \cite{Shephard:Andersen:2009}, the log-price
process $Y_t$ is serially uncorrelated although potentially dependent,
provided that $V_t$ and $\tau_t$ are independent.

Later authors, e.g. Back \cite{back:1991} chose to model the log-price
process as a semimartingale, with increments of the martingale
component modelled as a product process:
\begin{eqnarray}
  Y_t &=& Y_0 + A_t + M_t \nonumber \\
  M_t - M_{t-1} &=& X_t = \sigma_t Z_t
  \label{eq:rtght}
\end{eqnarray}
where $A_t$ is a finite variation process and $M_t$ is a martingale
and hence $Y_t$ is a semimartingale. $\sigma_t$ is non-negative and
$Z_t$ is an iid process with 0 mean and unit variance. A convenient
choice of $\sigma_t$ is
\begin{equation}
  \label{eq:rfht}
  \log\sigma_t = \sum_{l \in \mathbb Z} \psi_l \eta_{t-l}  
\end{equation}
where $\{\psi_l\}_{l \in \mathbb Z}$ is a sequence of real numbers
with at least one non-zero element,
$\{\eta_t\}_{t \in \mathbb Z}$ is an iid sequence of random
variables. By Kolmogorov's 3-series theorem, the infinite series above
converges if and only if $\sum_{l \in \mathbb Z} \psi_l^2 < \infty$,
and in this case $\{\sigma_{t}\}$ is stationary.
In particular, if $\{\eta_i\}_{i \in \mathbb Z}$ is normally
distributed with zero mean, $\{\sigma_t\}_{t \in \mathbb Z}$ has
log-normal marginal distributions.

If, however, the sequences $\{Z_t\}_{t \in \mathbb Z}$ or
$\{\sigma_t\}_{t \in \mathbb Z}$ are regularly varying with index
$\alpha$ and some additional conditions are satisfied, $X_t$ is also
regularly varying with the same index. Specifically, if $Z_t$ is
regularly varying and $\sigma_t$ has a lighter tail, the conclusion
follows from Breiman's lemma. See \S\ref{sec:case1} of Jan\ss en et
al. i.e. chapter \ref{ch:bernoulli} of this thesis for more details.

% If, instead, $\sigma_t$ has a heavier tail than $Z_t$, then $X_{i,t}$
% is regularly varying assuming the following conditions:
% \begin{itemize}
% \item $e^\eta$ is regularly varying with index $\alpha$, where $\eta$
%   is a generic element of the iid sequence
%   $\{\eta_t\}_{t \in \mathbb Z}$, and
%   \begin{equation}
%     \label{eq:dfhty}
%     \E e^{\alpha \eta} = \infty, \quad\text{ or }
%     {\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)} \to c \in (0, \infty)
%   \end{equation}
%     where $\eta, \eta_1$ and $\eta_2$ are generic elements of the
%     iid sequence $\{\eta_t\}_{t \in \mathbb Z}$.
% \end{itemize}
% Note the only finite limit for
% ${\P(\eta_1 + \eta_2 > x) \over \P(\eta_1 > x)}$ is
% $\E e^{\alpha \eta}$. Hence the two conditions of \eqref{eq:dfhty}
% are mutually exclusive. In chapter \ref{ch:bernoulli} we will see that
% these two conditions lead to rather different behavior for
% eigenvalues of the sample covariance matrix of stochastic volatility
% processes of the form \eqref{eq:rfht}.

There are a few advantages in using stochastic volatility models. For
one thing, they are among the simplest models allowing for conditional
heteroscedasticity (cf. Andersen et
al. \cite{andersen:davis:kreiss:mikosch:2009}); for another thing,
they greatly improve the accuracy of predicting future volatilities
by taking advantage of high frequency data. Specifically, for the
semimartingale $Y_t$ we have
\begin{equation}
  \label{eq:jieg}
  [Y_{\delta}]_t
  =
  \sum_{j=1}^{\floor{t/\delta}}
  (Y_{j \delta} - Y_{(j-1)\delta})^2
  \overset{P}{\to} [Y]_t,
  \quad
  \delta \to 0.
\end{equation}
The object $[Y_\delta]_t$ is called the realized quadratic variation
in time series literature.
The convergence in probability follows directly from the definition of
quadratic variation.
%% When $\E Z_t = 0$, $[Y]_t = [M]_t$ 
Furthermore, Jacod \cite{jacod:1994} and
Barndorff-Nielsen and Shephard \cite{barndorff:shephard:2002} proved a
central limit theorem:
\[
  {[Y_\delta]_t - [Y]_t \over \sqrt{2 \delta \int_0^t \sigma_s^2 ds}}
  \overset{d}{\to} N(0, 1),
  \quad \delta \to 0
\]
Moreover, the structure $X_t = M_{t} - M_{t-1} = \sigma_t Z_t$ gives
$[Y_\delta]_t \overset{P}{\to} \int_0^t \sigma_s^2 ds$. Meanwhile,
It\^o's formula gives, for semimartingale $Y_t$,
\begin{eqnarray}
  Y_t^2 &=& [Y]_t + 2 \int_0^t Y_s d Y_s \nonumber \\
  &=& [Y]_t + 2 \int_0^t Y_s dA_s + 2 \int_0^t Y_s dM_s \nonumber \\
  \E Y_t^2 &=& \E [Y]_t + 2 \E \left( \int_0^t Y_s dA_s \right) \nonumber\\
  &\approx& \E [Y]_t, \quad \text{when $t$ is small} \label{eq:hytre}
\end{eqnarray}
\eqref{eq:hytre} and \eqref{eq:jieg} show that $\E Y_t^2$, or in other
words, the forecast of future squared return can be obtained as
$\E [Y_\delta]_t$, i.e. the forecast of future realized quadratic
variation.

So far we have been discussing univariate stochastic volatility
models. In a multivariate context, we may model a $p$-variate process
as
\begin{eqnarray*}
  X_{i,t} &=& \sigma_{i,t} Z_{i,t}, \quad i = 1,2,\dots,p\\
  \log \sigma_{i, t}
  &=&
  \sum_{k,l \in \mathbb Z} \psi_{k,l} \eta_{i-k, t-l}
\end{eqnarray*}
where $Z_{i,t}$ is an iid field of random variables, and the
coefficients $\{\psi_{k,l}\}$ are real and satisfy
$\sum_{k, l \in \mathbb Z} |\psi_{k,l}| < \infty$.
% Since past realized quadratic variations are observable,
% forcast of future $[Y_\delta]_t$ can be obtained using e.g. an ARFIMA
% model.

@@

On the other hand, we have also seen that, when the left and the right
tails have the same indices, investor preference over the equity has
more sophisticated variations in the parameters' space including the
tail parameters of the equity, the interest rate, the investor's risk
appetite as captured by his utility function, and his threshold of
disappointment.
